---
title: "CTMax Report"
author: "Hallie Arno"
format: 
  html:
    code-fold: true
editor: source
---

```{r libraries}
#| message: false
#| warning: false

library(tidyverse)
library(data.table)
library(parsnip)
library(ranger)
library(tidymodels)
library(xgboost)
library(workflows)
library(stringi)
library(lubridate)
library(plotly)
library(kernelshap)
library(shapviz)

set.seed(12)

plinkPath <- "/opt/miniconda3/bin/plink2"

```


# Experiments and Linear Models

For detailed methods, see: <https://docs.google.com/document/d/17h9AMruVNdF4moGN1OnuQsRiXw9gWWQkvINjjDFYk-I/edit?usp=sharing>

```{r uploadMetadata}
#| echo: false

ambient <- read.csv("ctMaxReportMetadata/dailyAmbient.csv") %>% 
  rename("DailyTemp" = "TempC") %>% 
  select(-c(Date))

metadata <- fread("ctMaxReportMetadata/HallieCTMaxDatasheet.csv") %>% 
  select(c(Vial, Tank, Weight, Temp_C, Date)) %>% 
  rename("SampleID" = "Vial") %>% 
  mutate(Day = lubridate::yday(Date)-284) %>% 
  merge(ambient, by = "Day")


meta <- fread("ctMaxReportMetadata/CtMaxFish.fam") %>% 
  rename("FID"="V1",
         "IID" = "V2") %>% 
  select(c(FID, IID)) %>% 
   mutate(SampleIDStart = stri_extract(IID, regex = "(?<=_i5.).*")) %>% 
  mutate(SampleID = sub("\\.realigned.bam.*", "", SampleIDStart)) %>% 
  mutate(Population = gsub('[[:digit:]]+', '', SampleID)) %>% 
  select(-SampleIDStart) %>% 
  merge(metadata) %>% 
  filter(IID != "NS.X0112.002.UDP0201_i7---UDP0201_i5.TRB104.realigned.bam") #come back to later

duplicates <- meta %>% 
  select(FID, IID) %>% 
  group_by(IID) %>% 
  summarize(n = n()) %>% 
  filter(n > 1)
meta <- meta %>% 
   filter(!IID %in% duplicates$IID)

binFunction <- function(PC1, PC2) {
  if (PC1 > 0) {
    return("Terrenceville (Brown Trout)")
  } else if (PC1 < 0 & PC2 > 0) {
    return("Tailrace (Hybrid Swarm)")
  } else if (PC1 < 0 & PC2 < 0) {
    return("Aquabounty (Aquaculture)")
  }
}

#Fix so that group is based on genetic group instead of guessing

metaPCA <- fread("ctMaxReportMetadata/CtMaxFish.eigenvec") %>% 
  rename("FID" = "V1",
         "IID" = "V2",
         "PC1" = "V3",
         "PC2" = "V4",
         "PC3" = "V5") %>% 
  select(c(FID, IID, PC1, PC2, PC3)) %>% 
  merge(meta) %>% 
  rowwise() %>% 
  mutate(Group = binFunction(PC1, PC2)) %>% 
  filter(!IID %in% duplicates$IID)
```

```{r boxplot}
#| warning: false
#| echo: false

ggplot(metaPCA, aes(x = Group, y = Temp_C)) + 
  geom_boxplot(colour = "darkblue") + 
  theme_bw() +
  ylab("Critical Thermal Maximum (C)")
```
Fish from the wild river had the highest CTMax. Aquaculture fish had the lowest CTMax, while Terrenceville Trout were intermediate. Differences between groups are statistically significant (see models below).

## GLMs
Questions: 

<b>
1: Does group (ancestry) impact CTMax? 

2: Does ambient temperature impact CTMax?

3: How much does weight account for CTMax? 
</b>

*GzLMs are used instead of an lm() because it is nuetral to the order of the terms.*

Variables: 

Group: Population (Tailrace, Terrenceville, Aquabounty)

Weight: Wight in grams after CTMax Trial.

Daily Temp: Ambient temperature that fish were kept at the morning of the CTMax trial. All fish were at the same ambient temperature. 

*Reviewers may ask for mixed models using trial date or tank as a random factor, I am working on these.*


### Model with both groups, including weight, daily ambient temperature, and interaction term.
```{r omnibusModel}
model1 <- glm(Temp_C~Group + Weight + DailyTemp + Group*Weight + Group*DailyTemp + Weight*DailyTemp, data = metaPCA, family = gaussian(link = "identity"))

```

::: {.panel-tabset}
#### ANOVA Table
```{r}
rmarkdown::paged_table(anova(model1)%>% select(-c(Df, `Resid. Df`)))

```
#### Effect Sizes (Summary)
```{r}
summary(model1)
```
#### Residual Plots
:::: {#fig-plots layout-ncol=2}
```{r}
plot(model1)
```
::::

:::

When including all populations and ambient temperature in the model, population is the most important factor in predicting CTMax. Weight alone is not a significant predictor overall (though note, it is correlated with both group and ambient daily temperature) but may be significant within group (see interaction term in summary). Ambient daily temperature is significant in the ANOVA, but has a small effect size and is not significant in the summary (this is likely because of the method used to detect significance in each; because there is only one temperature for all values that day there could be some weird results). 

### Model with aquaculture fish only, including weight, daily ambient temperature, and interaction term.


```{r aqcModel}
aqcMeta <- metaPCA %>% 
  filter(Group == "Aquabounty (Aquaculture)")

modelA <- glm(Temp_C~Weight + DailyTemp + Weight*DailyTemp, data = aqcMeta, family = gaussian(link = "identity"))
```

::: {.panel-tabset}
#### ANOVA Table
```{r}
rmarkdown::paged_table(anova(modelA)%>% select(-c(Df, `Resid. Df`)))

```
#### Effect Sizes (Summary)
```{r}
summary(modelA)
```

#### Residual Plots
:::: {#fig-plots layout-ncol=2}
```{r}
plot(modelA)
```
::::

:::

This model shows similar results to the omnibus model. In the ANOVA table, both weight and ambient temperature are significant. Weight has a much larger effect size in the summary. Again, these are correlated, so it may be difficult to pull apart the variance from each. 

### Model with aquaculture fish only, including weight, daily ambient temperature, and interaction term.

```{r wildModel}
wildMeta <- metaPCA %>% 
  filter(Group == "Tailrace (Hybrid Swarm)")

modelW <- glm(Temp_C~Weight + DailyTemp + Weight*DailyTemp, data = wildMeta, family = gaussian(link = "identity"))
```

::: {.panel-tabset}
#### ANOVA Table
```{r}
rmarkdown::paged_table(anova(modelW) %>% select(-c(Df, `Resid. Df`)))

```
#### Effect Sizes (Summary)
```{r}
summary(modelW)
```

#### Residual Plots
:::: {#fig-plots layout-ncol=2}
```{r}
plot(modelW)
```
::::

:::

For wild fish, ambient daily temperature has a larger effect size and significance than weight. Maybe this is because they were more uniform in weight from beginning to end of the trials? Or maybe this has something to do with adaptation to fluctuating temperatures in the wild. Perhaps something in the methylation data would provide insight. I think this result is really interesting, especially in light of recent research showing acclimation temperature is extremely important for CTMax in hatchery fish. 

### Summary Plots
::: {.panel-tabset}
#### Weight Distribution
```{r weights}
#| warning: false
#| echo: false
ggplot(metaPCA, aes(x = Weight, fill = Group, colour = Group)) + 
  geom_density(alpha = 0.3) + 
  theme_bw() + 
  ggtitle("Distribution of Weights")
```

#### Ambient Temp.
```{r ambientTemp}
#| warning: false
#| echo: false

ggplot(metaPCA, aes(x = Day, y = DailyTemp)) + 
  geom_point() + 
  theme_bw() + 
  ggtitle("Ambient Daily Temperature At Each Trial") + 
  xlab("Trial Day") + 
  ylab("Ambient Temperature (C)")
```

#### Weight vs. Ambient Temp
How Correlated are Daily Temp and Weight Really?
```{r weightVTemp}
#| warning: false
#| echo: false

ggplot(metaPCA, aes(y = Weight, x = DailyTemp, colour = Group)) + 
  geom_point() + 
  stat_smooth(method = "lm") + 
 # facet_wrap(.~Group, ncol = 1) +
  theme_bw() + 
  ggtitle("Ambient Daily Temperature Versus Weight") + 
  xlab("Ambient Temp (C)") + 
  ylab("Weight (g)")
```
:::

# Genetic Data
Population Structure\
*click and drag to zoom*
```{r pca}
#| warning: false
#| echo: false

plot <- ggplot(metaPCA, aes(x = PC1, y = PC2, label = SampleID, colour = Population)) + 
  geom_text(size = 2) + 
  theme_bw() 

ggplotly(plot)
```

*Some of the labels may be mistakes. Because of that, all models above use genetic group, not labelled group. Aquaculture fish are considered samples with PC1 & PC2 < 0.*

There are likely 3 salmon/ trout hybrids. I am still not sure why there are two groups of farmed fish (perhaps they are closely related family groups). Most TRB fish are fairly tighly clustered, suggesting similar levels of introgression across the population. I don't think there are any F1 aquaculture/ wild hybrids. However, there are a few fish in TRB cluster with higher than expected PC1 (such as TRB75 and TRB84). Could this be EU introgression? Hopefully running admixture with other aquaculture samples will provide insight. 

## Salmon GWAS
*The rest of this document uses only salmon samples (PC1 < 0).*

### PLINK GWAS
The first step was conducting a basic GWAS in PLINK. I ran the GWAS seperately between wild and aquaculture, since running them together resulted in SNPs highly differential between groups, not necessarily associated with thermal tolerance. Note these models use no covariates.

Using Bonferroni false discovery rate correction for p-values: 

```{r bonferroni}
#| warning: false
#| message: false
#| 
bonferroni <- .05/5510179


trbGWASbonferroni <- fread("ctMaxReportMetadata/TailraceGWASnoCovar.Temp_C.glm.linear") %>% 
  filter(P <= bonferroni) 
rmarkdown::paged_table(trbGWASbonferroni)

fffGWASbonferroni <- fread("ctMaxReportMetadata/AquabountyGWASnoCovar.Temp_C.glm.linear") %>% 
  filter(P <= bonferroni) 

rmarkdown::paged_table(fffGWASbonferroni)

```

No SNPs are significant in TRB using a bonferroni p-value correction. 3 are significant in the farmed fish. 

However, with a somewhat arbitrarily chosen alpha < 10^-4:
```{r 10e04}
#| warning: false
#| message: false

trbGWAS10e4 <- fread("ctMaxReportMetadata/TailraceGWASnoCovar.Temp_C.glm.linear") %>% 
  filter(P <= 0.0001) %>% 
  mutate(Group = "TRB")
rmarkdown::paged_table(trbGWAS10e4)

fffGWAS10e4 <- fread("ctMaxReportMetadata/AquabountyGWASnoCovar.Temp_C.glm.linear") %>% 
  filter(P <= 0.0001 ) %>% 
  mutate(Group = "F")

rmarkdown::paged_table(fffGWAS10e4)

```
There are many more significant aquaculture SNPs. 

```{r PLINKmanhattanPlots}
#| message: false
#| warning: false

chr <- read_csv("ctMaxReportMetadata/ssal_chromosme_conversion.csv") %>% 
  separate_wider_delim(ChrNum, delim = "a", names = c("Sp", "Number")) %>% 
  select(NC, Number) %>% 
  rename("CHROM" = "NC")



prepData <- function(df) {
  myfile <- as.data.table(df)
  #Make position numeric
  myfile[, POS := as.numeric(POS)]
  

  #This determines the max bp for each chromosome 
  data_cumu <- myfile[,.(maxBP=max(POS)),.(Number)];
  
  #This makes a dt with chr and BP to add for each
  setDT(data_cumu)[, cum_sum := shift(cumsum(as.numeric(maxBP)), fill = 0)] 
  
  #Merge and add BP from above
  new <- merge(myfile, data_cumu,  by = "Number")
  
  new[, BP := POS + cum_sum]
  
  
  #Bind together and remove unnecessary rows
  return(new[, c("maxBP", "cum_sum") := NULL])
  

}

combinedGWAS <- rbind(trbGWAS10e4, fffGWAS10e4) %>% 
  rename("CHROM" = `#CHROM`) %>% 
  merge(chr) %>% 
  select(-CHROM)
  
GWAS <- prepData(combinedGWAS)


center <- GWAS[,.(center=mean(BP)),.(Number)];


```

::: {.panel-tabset}
#### P-value
```{r pValManhattanPlot}


mp <-  ggplot() + 
   geom_point(data = GWAS, aes(x = BP, y = -log(P, base = 10), colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
  scale_x_continuous(
    label = center$Number,
    breaks = center$center) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(GWAS$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "-log(P)") +
  facet_wrap(~Group, ncol = 1)
 
mp
```

#### Beta
```{r BetaManhattanPlot}


mb <-  ggplot() + 
   geom_point(data = GWAS, aes(x = BP, y = BETA, colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
  scale_x_continuous(
    label = center$Number,
    breaks = center$center) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(GWAS$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "BETA") +
  facet_wrap(~Group, ncol = 1)
 
mb
```

:::

### Overlap Between Wild and Farmed
```{r}
overlap <- merge(trbGWAS10e4, fffGWAS10e4, by = "ID") %>% 
  select(c(ID, `#CHROM.x`, POS.x, BETA.x, BETA.y, P.x, P.y)) %>% 
  rename(
    "CHROM" = `#CHROM.x`,
    "POS" = "POS.x"
  ) %>% 
  merge(chr, by = "CHROM")
rmarkdown::paged_table(overlap)
```
Just one of these is in a Gene (LOC106565882):

```{r}
write.table(overlap %>%  
              mutate(POSend = POS) %>% 
              select(c(CHROM, POS, POSend, ID)),
  "ctMaxReportMetadata/10e4Overlap.bed", col.names = F, row.names = F, sep = "\t", quote = F)


system("bedtools intersect -b ctMaxReportMetadata/10e4Overlap.bed -a ctMaxReportMetadata/Ssal3_1.genes.unique.bed -wb >  ctMaxReportMetadata/10e4OverlapGenes.tsv")
rmarkdown::paged_table(fread("ctMaxReportMetadata/10e4OverlapGenes.tsv"))
```


The p < 10e4 subset of SNPs is sufficient to run random forest models. 

## Machine Learning Models

To determine SNPs most predictive of CTMax, I tried two machine learning models: random forest and gradient boosting. For both, I started with SNPs from the GWAS above (in part to just cut down on the processing power needed).

### Random Forest

For the random forest models, I iteratively removed the SNPs in the lower half of prediction power until the rsquared stopped increasing. The final SNPs I was left with are run here. 


```{r RFDataPrep}
#| message: false

tailrace <- metaPCA %>% 
  filter(Group == "Tailrace (Hybrid Swarm)") %>% 
  filter(!is.na(Temp_C)) %>% 
  distinct()

indTRB <- sample(c(TRUE, FALSE), length(unique(tailrace$IID)), replace = TRUE, prob = c(0.6, 0.4))

trainTRB <- tailrace[indTRB, ]
testTRB <- tailrace[!indTRB, ]

write.table(trainTRB %>% select(FID, IID) %>% distinct(), 
            file = "ctMaxReportMetadata/TrbTrainCTmax.tsv", 
            col.names = F, 
            row.names = F, 
            sep = "\t", 
            quote = F)

write.table(testTRB %>% select(FID, IID) %>% distinct(), 
            file = "ctMaxReportMetadata/TrbTestCTmax.tsv", 
            col.names = F, 
            row.names = F, 
            sep = "\t", 
            quote = F)

## Aqc train/ test
aquabounty <- metaPCA %>% 
  filter(Group == "Aquabounty (Aquaculture)")

indF <- sample(c(TRUE, FALSE), length(unique(aquabounty$IID)), replace = TRUE, prob = c(0.6, 0.4))

trainF <- aquabounty[indF, ]
testF<- aquabounty[!indF, ]

write.table(trainF %>% select(FID, IID) %>% distinct(), 
            file = "ctMaxReportMetadata/FTrainCTmax.tsv", 
            col.names = F, 
            row.names = F, 
            sep = "\t", 
            quote = F)

write.table(testF %>% select(FID, IID) %>% distinct(), 
            file = "ctMaxReportMetadata/FTestCTmax.tsv", 
            col.names = F, 
            row.names = F, 
            sep = "\t", 
            quote = F)
```


```{r RFPLINK}
#| message: false
#| output: false

system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/TrbTrainCTmax.tsv --extract ctMaxReportMetadata/ThSnRfSNPsTRB.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTrainTRB.ThSn --make-bed"))
system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/TrbTestCTmax.tsv --extract ctMaxReportMetadata/ThSnRfSNPsTRB.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTestTRB.ThSn --make-bed"))


system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/FTrainCTmax.tsv --extract ctMaxReportMetadata/SxFrRfSNPsF.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTrainF.SxFr  --make-bed" ))
system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/FTestCTmax.tsv --extract ctMaxReportMetadata/SxFrRfSNPsF.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTestF.SxFr  --make-bed" ))


```


```{r randomForest}
#| message: false
# Need to change names! farm is 256!

rf_generic <- rand_forest() %>% 
  set_mode("regression") %>% 
  set_engine("ranger", importance = "permutation") #need to look more into importance type


trainfamTRBThSn <- fread("ctMaxReportMetadata/ctmaxTrainTRB.ThSn.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
  inner_join(meta) %>% 
  distinct()


dataTrainTRBThSn <- fread("ctMaxReportMetadata/ctmaxTrainTRB.ThSn.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

trbTrainImpThSn <- data.frame(apply(dataTrainTRBThSn, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))) 

TRBForRFThSn <- na.omit(cbind(trbTrainImpThSn, trainfamTRBThSn$Temp_C)) %>% 
  rename( "CtMax" = "trainfamTRBThSn$Temp_C")

trbRecipeThSn <- recipe(CtMax ~ ., data = TRBForRFThSn) 
 
trbWfThSn <- workflow(trbRecipeThSn, rf_generic)

trbFitThSn <- fit(trbWfThSn, TRBForRFThSn) #rsq = 0.7289942

trbRfImpThSn <- data.frame(trbFitThSn$fit$fit$fit$variable.importance) %>% 
  arrange(desc(trbFitThSn.fit.fit.fit.variable.importance)) %>% 
  rownames_to_column() %>% 
  rename("Feature" = "rowname",
         "RFImportance" = "trbFitThSn.fit.fit.fit.variable.importance") %>% 
  mutate(Group = "TRB")
#### Repeat for farm


trainfamFSxFr <- fread("ctMaxReportMetadata/ctmaxTrainF.SxFr.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
  inner_join(meta) %>% 
  distinct()


dataTrainFSxFr <- fread("ctMaxReportMetadata/ctmaxTrainF.SxFr.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

fTrainImpSxFr <- data.frame(apply(dataTrainFSxFr, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))) 

fForRFSxFr <- na.omit(cbind(fTrainImpSxFr, trainfamFSxFr$Temp_C)) %>% 
  rename( "CtMax" = "trainfamFSxFr$Temp_C")


fRecipeSxFr <- recipe(CtMax ~ ., data = fForRFSxFr) 
 
fWfSxFr <- workflow(fRecipeSxFr, rf_generic)

fFitSxFr <- fit(fWfSxFr, fForRFSxFr) #rsq = 0.7238919 

fRfImpSxFr <- data.frame(fFitSxFr$fit$fit$fit$variable.importance) %>% 
  arrange(desc(fFitSxFr.fit.fit.fit.variable.importance)) %>% 
  rownames_to_column() %>% 
  rename("Feature" = "rowname",
         "RFImportance" = "fFitSxFr.fit.fit.fit.variable.importance") %>% 
  mutate(Group = "F")


combinedRF <- rbind(fRfImpSxFr, trbRfImpThSn) %>%  
  separate_wider_delim(Feature, delim = ".1_", names = c("NC", "POS")) %>% 
  separate_wider_delim(POS, delim = "_", names = c("POS", "N")) %>%
  mutate(NC = paste0(NC, ".1")) %>% 
  mutate(ID = paste0(NC, "_", POS)) %>% 
  mutate(Feature = paste0(ID, "_", N)) %>% 
  merge(chr, by.x = "NC", by.y = "CHROM") %>% 
  select(c(Number, POS, Feature, Group, RFImportance)) %>% 
  arrange(desc(RFImportance)) %>% 
  mutate(Number = as.numeric(Number))
```

::: {.panel-tabset}

#### SNPs
```{r}
rmarkdown::paged_table(combinedRF)
```

#### Aqc Summary
```{r}
fFitSxFr
```

#### Wild Summary
```{r}
trbFitThSn 
```

:::

#### RF Test
```{r}
#| message: false
trbTestFam <- fread("ctMaxReportMetadata/ctmaxTrainTRB.ThSn.fam") %>% 
  select(V2) %>% 
  rename("IID" = "V2") %>% 
  left_join(metaPCA)

trbTestSNPs <- fread("ctMaxReportMetadata/ctmaxTrainTRB.ThSn.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

trbTest <- cbind(trbTestSNPs, trbTestFam$Temp_C)

predictionTRB <- predict(trbFitThSn, new_data = trbTest) %>% 
  cbind(trbTestFam$Temp_C, trbTestFam$SampleID) %>% 
  rename("Prediction" = ".pred",
         "Actual" = "trbTestFam$Temp_C",
         "SampleID" = "trbTestFam$SampleID") %>% 
  mutate(Group = "TRB")

### Farm

fTestFam <- fread("ctMaxReportMetadata/ctmaxTrainF.SxFr.fam") %>% 
  select(V2) %>% 
  rename("IID" = "V2") %>% 
  left_join(metaPCA)

fTestSNPs <- fread("ctMaxReportMetadata/ctmaxTrainF.SxFr.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

fTest <- cbind(fTestSNPs, fTestFam$Temp_C)

predictionF <- predict(fFitSxFr, new_data = fTest) %>% 
  cbind(fTestFam$Temp_C, fTestFam$SampleID) %>% 
  rename("Prediction" = ".pred",
         "Actual" = "fTestFam$Temp_C",
         "SampleID" = "fTestFam$SampleID") %>% 
  mutate(Group = "F") 

predPlot <- rbind(predictionTRB, predictionF) %>% 
  ggplot(aes(x = Prediction, y = Actual, colour = Group, label = SampleID)) +
  geom_point() + 
  theme_bw() + 
  labs(title = "Predicted vs. Actual CTMax from Random Forest Models")

predPlot
```


#### Manhattan Plot of Random Forest results

```{r}
rfMP <- prepData(combinedRF %>%  arrange(Number)) 

center <- rfMP[,.(center=mean(BP)),.(Number)];


mpRF <-  ggplot() + 
   geom_point(data = rfMP, aes(x = BP, y = RFImportance, shape = as.factor(Group), colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
   scale_x_continuous(
     label = center$Number,
     breaks = center$center
     ) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(rfMP$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "Random Forest Importance") +
  facet_wrap(~Group, ncol = 1)
 
mpRF

```
There are no SNP overlaps between wild and farmed. 
```{r}
rmarkdown::paged_table(merge(fRfImpSxFr, trbRfImpThSn, by = "Feature"))
```

Does RF Importance correlate with a low p-value? 
```{r}
#| message: false
#| warning: false
plinkRfTRB<- trbGWAS10e4 %>% 
  mutate(Feature = paste0(ID, "_", REF)) %>% 
  merge(trbRfImpThSn)

plinkRfF<- fffGWAS10e4 %>% 
  mutate(Feature = paste0(ID, "_", REF)) %>% 
  merge(fRfImpSxFr)

plinkRfTrbF <- rbind(plinkRfTRB, plinkRfF)

ggplot(plinkRfTrbF, aes(x = RFImportance, y = -log(P, base = 10), 
                          colour = Group)) +
  geom_point() +
  theme_bw() +
  stat_smooth(method = "lm")
```

#### Bedtools intersect genes: 
```{r}
finalF <- fRfImpSxFr %>% 
  separate_wider_delim(Feature, delim = ".1_", names = c("NC", "POS")) %>% 
  separate_wider_delim(POS, delim = "_", names = c("POS", "N")) %>%
  mutate(NC = paste0(NC, ".1")) %>% 
  mutate(ID = paste0(NC, "_", POS)) %>% 
  merge(chr) %>% 
  mutate(POSend = POS) 


finalTRB <- trbRfImpThSn %>% 
  separate_wider_delim(Feature, delim = ".1_", names = c("NC", "POS")) %>% 
  separate_wider_delim(POS, delim = "_", names = c("POS", "N")) %>%
  mutate(NC = paste0(NC, ".1")) %>% 
  mutate(ID = paste0(NC, "_", POS)) %>% 
  merge(chr) %>% 
  mutate(POSend = POS) 



write.table(finalTRB %>%  
              distinct() %>% 
              select(c(NC, POS, POSend, ID)),
  "ctMaxReportMetadata/FinalTRB.bed", col.names = F, row.names = F, sep = "\t", quote = F)

write.table(finalF %>%  
              distinct() %>% 
              select(c(NC, POS, POSend, ID)),
  "ctMaxReportMetadata/FinalF.bed", col.names = F, row.names = F, sep = "\t", quote = F)


system("bedtools intersect -b ctMaxReportMetadata/FinalTRB.bed -a ctMaxReportMetadata/Ssal3_1.genes.unique.bed -wb >  ctMaxReportMetadata/RF.Small.TRB.genes.tsv")
system("bedtools intersect -b ctMaxReportMetadata/FinalF.bed -a ctMaxReportMetadata/Ssal3_1.genes.unique.bed -wb >  ctMaxReportMetadata/RF.Small.F.genes.tsv")

trbGene <- fread("ctMaxReportMetadata/RF.Small.TRB.genes.tsv") %>% 
  rename("CHROM" = "V1",
         "POS" = "V6", 
         "ID" = "V8",
         "Gene" = "V4") %>%
  distinct() %>% 
  mutate(Group = "TRB") %>% 
  merge(plinkRfTRB, all.y = FALSE) %>% 
  merge(chr, by = "CHROM") %>% 
  select(c(Number, POS, Gene, Feature, Group, BETA, P, RFImportance))
  
fGene <- fread("ctMaxReportMetadata/RF.Small.F.genes.tsv") %>% 
  rename("CHROM" = "V1",
         "POS" = "V6", 
         "ID" = "V8",
          "Gene" = "V4") %>%
  distinct() %>% 
  mutate(Group = "F") %>% 
  merge(plinkRfF, all.y = FALSE) %>% 
  merge(chr, by = "CHROM") %>% 
  select(c(Number, POS, Gene, Feature, Group, BETA, P, RFImportance))
```

Tailrace Top RF SNPs in Genes: 
```{r}
rmarkdown::paged_table(trbGene)
```



Aquaculture: Top RF SNPs in Genes: 
```{r}
rmarkdown::paged_table(fGene)
```

*Neither of these lists of genes create pathways in ShinyGO (too few genes for TRB)*


There are no gene overlaps between wild and aquaculture: 
```{r}
geneOverlap <- merge(trbGene, fGene, by = "Gene") #no gene overlaps
rmarkdown::paged_table(geneOverlap)
```



### Gradient Boosting


For gradient boosting, I used a large subset of SNPs (not the reduced RF results).
```{r xgBoostDataPrepTables}
#| message: false


write.table(trbGWAS10e4 %>%  
              mutate(POSend = POS) %>% 
              select(`#CHROM`, POS, POSend, ID) %>%
              distinct(),
  "ctMaxReportMetadata/TRBctmaxGWASsnps.bed", col.names = F, row.names = F, sep = "\t", quote = F)


write.table(fffGWAS10e4 %>%  
              filter(P < 0.00001) %>% 
                 mutate(POSend = POS) %>% 
              select(`#CHROM`, POS, POSend, ID) %>%
              distinct(),
  "ctMaxReportMetadata/FFFctmaxGWASsnps.bed", col.names = F, row.names = F, sep = "\t", quote = F)
```


```{r xgBoostPLINK}
#| output: false
# TRB
system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/TrbTrainCTmax.tsv --extract ctMaxReportMetadata/TRBctmaxGWASsnps.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTrainTRB  --make-bed" ))

system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/TrbTestCTmax.tsv --extract ctMaxReportMetadata/TRBctmaxGWASsnps.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTestTRB --make-bed" ))

# Aqc

system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/FTrainCTmax.tsv --extract ctMaxReportMetadata/FFFctmaxGWASsnps.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTrainFFF --mind 0.2 --make-bed" ))

system(paste(plinkPath, "--bfile ctMaxReportMetadata/suspectedSalmon --keep ctMaxReportMetadata/FTestCTmax.tsv --extract ctMaxReportMetadata/FFFctmaxGWASsnps.bed --allow-extra-chr --chr-set 29 --recode A --out ctMaxReportMetadata/ctmaxTestFFF --mind 0.2 --make-bed" ))
```


```{r xgBoostDataPrep}
#| message: false
#| 
trainfamTRB <- fread("ctMaxReportMetadata/ctmaxTrainTRB.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
   inner_join(metaPCA) %>% 
   distinct()


dataTrainTRB <- fread("ctMaxReportMetadata/ctmaxTrainTRB.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)
#genos.dose.train.rda <- fread(paste0("NA.train.raw"), sep = " ") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

trbTrainImp<- data.frame(apply(dataTrainTRB, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))) 

TRBForRF <- na.omit(cbind(trbTrainImp, trainfamTRB$Temp_C)) %>% 
  rename( "CtMax" = "trainfamTRB$Temp_C")


###########
## TEST ##
###########

testfamTRB <- fread("ctMaxReportMetadata/ctmaxTestTRB.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
  inner_join(metaPCA) 


dataTestTRB <- fread("ctMaxReportMetadata/ctmaxTestTRB.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)
#genos.dose.train.rda <- fread(paste0("NA.train.raw"), sep = " ") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

trbTestImp<- data.frame(apply(dataTestTRB, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x)))))))

trbTest <- cbind(trbTestImp, testfamTRB$Temp_C) %>% 
  rename("Temp_C" = "testfamTRB$Temp_C")

##############
### Farm ####
#############

trainfamF <- fread("ctMaxReportMetadata/ctmaxTrainFFF.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
  inner_join(aquabounty) %>% 
  distinct()


dataTrainF <- fread("ctMaxReportMetadata/ctmaxTrainFFF.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)
#genos.dose.train.rda <- fread(paste0("NA.train.raw"), sep = " ") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

fffTrainImp<- data.frame(apply(dataTrainF, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))) 

FForRF <- na.omit(cbind(fffTrainImp, trainfamF$Temp_C)) %>% 
  rename("CtMax"= "trainfamF$Temp_C")


###########
## TEST ##
###########

testfamFFF <- fread("ctMaxReportMetadata/ctmaxTestFFF.fam") %>% 
  select(V1, V2) %>% 
  rename("FID" = "V1",
         "IID" = "V2") %>% 
  inner_join(aquabounty) %>% 
  distinct()


dataTestF <- fread("ctMaxReportMetadata/ctmaxTestFFF.raw", sep = "\t") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)
#genos.dose.train.rda <- fread(paste0("NA.train.raw"), sep = " ") %>%  select (-FID, -IID, -PAT, -MAT, -SEX, -PHENOTYPE)

fffTestImp<- data.frame(apply(dataTestF, 2, function(x) replace(x, is.na(x), as.numeric(names(which.max(table(x))))))) 

fTest <- cbind(fffTestImp, testfamFFF$Temp_C) %>% 
  rename("Temp_C" = "testfamFFF$Temp_C")

```

```{r xgBoostModel}
#| message: false
#
trbRecipe <- recipe(CtMax ~ ., data = TRBForRF) 
 ## need to giure out params!!


TRBxgboostModel <- boost_tree(trees = 1000, tree_depth = 2, learn_rate = 0.1) %>%
  set_mode("regression") %>% 
  set_engine("xgboost")

trbXgboostWf <- workflow(trbRecipe, TRBxgboostModel)

trbXgFit <- fit(trbXgboostWf, TRBForRF)
 
trbXgImp <- xgb.importance(trbXgFit$fit$fit$fit) %>% 
  mutate(ID = substr(Feature,1,nchar(Feature)-2)) %>% 
   arrange(desc(Gain)) %>% 
#  mutate(rank = row_number()) %>% 
  mutate(Group = "TRB")



fRecipe <- recipe(CtMax ~ ., data = FForRF) 
 ## need to giure out params!!


FxgboostModel <- boost_tree(trees = 1000, tree_depth = 2, learn_rate = 0.1) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")

fXgboostWf <- workflow(fRecipe, FxgboostModel)

fXgFit <- fit(fXgboostWf, FForRF)
 
fXgImp <- xgb.importance(fXgFit$fit$fit$fit) %>% 
  mutate(ID = substr(Feature,1,nchar(Feature)-2)) %>% 
   arrange(desc(Gain)) %>% 
 # mutate(rank = row_number()) %>% 
  mutate(Group = "F")

xgBoostImp <- rbind(trbXgImp, fXgImp) %>% 
   separate_wider_delim(ID, delim = ".1_", names = c("NC", "POS")) %>% 
  #separate_wider_delim(POS, delim = "_", names = c("POS", "N")) %>%
  mutate(NC = paste0(NC, ".1")) %>% 
 # mutate(ID = paste0(NC, "_", POS)) %>% 
  merge(chr, by.x = "NC", by.y = "CHROM")



```
#### XGBoost Test
```{r xgboostTest}

predictionTRB <- predict(trbXgFit, new_data = trbTest) %>% 
  cbind(testfamTRB$Temp_C, testfamTRB$SampleID) %>% 
  rename("Prediction" = ".pred",
         "Actual" = "testfamTRB$Temp_C",
         "SampleID" = "testfamTRB$SampleID") %>% 
  mutate(Group = "TRB") 


predictionF <- predict(fXgFit, new_data = fTest) %>% 
  cbind(testfamFFF$Temp_C, testfamFFF$SampleID) %>% 
  rename("Prediction" = ".pred",
         "Actual" = "testfamFFF$Temp_C",
         "SampleID" = "testfamFFF$SampleID") %>% 
  mutate(Group = "F") 

xgboostPred <- rbind(predictionTRB, predictionF) %>% 
  ggplot(aes(x = Prediction, y = Actual, colour = Group)) +
  geom_point() + 
  theme_bw() + 
  labs(title = "Gradient Boosting Predicted vs. Actual CTMax")

xgboostPred
```


#### Manhattan Plot of xgBoost Results

xgboost model importance only provides information for SNPs used in tree splits (most important variables), so I did not determine an importance cutoff, but plotted all information provided from the model.



::: {.panel-tabset}

##### Gain
```{r gainMP}
xgBoostMP <- prepData(xgBoostImp)

center <- xgBoostMP[,.(center=mean(BP)),.(Number)];


mpXG <-  ggplot() + 
   geom_point(data = xgBoostMP, aes(x = BP, y = Gain, colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
  scale_x_continuous(
    label = center$Number,
    breaks = center$center) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(xgBoostMP$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "Gain") +
  facet_wrap(~Group, ncol = 1)
 
mpXG



```

##### Cover
```{r coverMP}
mpXC <-  ggplot() + 
   geom_point(data = xgBoostMP, aes(x = BP, y = Cover, colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
  scale_x_continuous(
    label = center$Number,
    breaks = center$center) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(xgBoostMP$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "Cover") +
  facet_wrap(~Group, ncol = 1)
 
mpXC

```

##### Frequency
```{r FreqMP}
mpXF <-  ggplot() + 
   geom_point(data = xgBoostMP, aes(x = BP, y = Frequency, colour = as.factor(Number)),  alpha = 1, size = 2) +
  theme_bw() + 
  scale_x_continuous(
    label = center$Number,
    breaks = center$center) +
  scale_colour_manual(values = rep(
    c("steelblue", "gray40"),
    unique(length(xgBoostMP$Number))
  )) +  
  theme(
    legend.position = "none",
    #panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 60, size = 8, vjust = 0.5) 
  ) +
  labs(x = "Position", y = "Frequency") +
  facet_wrap(~Group, ncol = 1)
 
mpXF

```

:::

#### Compare TRB and F:
```{r TRBvFxgboost}
rmarkdown::paged_table(merge(fXgImp, trbXgImp, by = "Feature"))
```

No overlap. 

#### Bedtools intersect with genes:

Compare TRB and F by gene: 
```{r trbFGeneOverlapXg}
#| message: false

trbForGeneXG <- xgBoostImp %>% 
  filter(Group == "TRB") %>% 
  mutate(ID = paste0(NC, "_", POS))

fForGeneXG <- xgBoostImp %>% 
  filter(Group == "F") %>% 
  mutate(ID = paste0(NC, "_", POS))

write.table(trbForGeneXG %>%  
              mutate(POSend = POS) %>% 
              select(c(NC, POS, POSend, ID)),
  "ctMaxReportMetadata/xgBoostTRB.bed", col.names = F, row.names = F, sep = "\t", quote = F)

write.table(fForGeneXG %>%  
              mutate(POSend = POS) %>% 
              select(c(NC, POS, POSend, ID)),
  "ctMaxReportMetadata/xgBoostF.bed", col.names = F, row.names = F, sep = "\t", quote = F)


system("bedtools intersect -b ctMaxReportMetadata/xgBoostTRB.bed -a ctMaxReportMetadata/Ssal3_1.genes.unique.bed -wb >  ctMaxReportMetadata/xgBoostTRBGene.tsv")

system("bedtools intersect -b ctMaxReportMetadata/xgBoostF.bed -a ctMaxReportMetadata/Ssal3_1.genes.unique.bed -wb >  ctMaxReportMetadata/xgBoostFGene.tsv")


rmarkdown::paged_table(merge(fread("ctMaxReportMetadata/xgBoostTRBGene.tsv"), fread("ctMaxReportMetadata/xgBoostFGene.tsv"), by = "V4") %>% 
  select(V4))

```


ShinyGO for each list seperately reveals 1 pathway for TRB (Arginine and proline metabolism) and five for aquaculture: Biosynthesis of unsaturated fatty acids, Fatty acid metabolism, PPAR signaling pathway, MRNA surveillance pathway, Tight junction. However, all except tight junction only have two genes in the pathway. 

#### Compare to RF results:

For now, I will use gain to compare to random forest, which is generally considered the best indicator of gradient boosting importance (though I will read more about this!)


There is a moderate amount of SNP overlap: 
```{r xgBoostToRF}

merge(xgBoostImp, combinedRF, by = c("Feature", "Group", "Number", "POS")) %>% 
  ggplot(aes(x = Gain, y = RFImportance, color = Group)) +
  geom_point() + 
  theme_bw()
```
RF + xgboost **distinct** genes overlap for TRB: 
```{r}
##trb
rmarkdown::paged_table(merge(fread("ctMaxReportMetadata/xgBoostTRBGene.tsv"), fread("ctMaxReportMetadata/RF.Small.TRB.genes.tsv"), by = "V4") %>% 
  select(V4) %>%
  distinct())

```
For aquaculture: 
```{r}
##farm
rmarkdown::paged_table(merge(fread("ctMaxReportMetadata/xgBoostFGene.tsv"), fread("ctMaxReportMetadata/RF.Small.F.genes.tsv"), by = "V4") %>% 
  select(V4) %>%
  distinct())
```

All of these overlaps change slighlty depending on RF/ XGBoost run. I think model tuning would make the results more stable, though there may always be some stochasticity.

```{r xgBoostEpistasis}
#| output: false
#| eval: false

# SHAP analysis
df_explainTRB <- TRBForRF[1:96, ]

X_predTRB <- bake(  # Goes to xgboost:::predict.xgb.Booster()
  prep(trbRecipe), 
  has_role("predictor"),
  new_data = df_explainTRB,
  composition = "matrix"
)

stopifnot(colnames(X_predTRB) %in% colnames(df_explainTRB))

shap_valuesTRB <- extract_fit_engine(trbXgFit) |> 
  shapviz(X_pred = X_predTRB, X = df_explainTRB, interactions = TRUE)

# SHAP importance
# shap_values |> 
#   sv_importance(show_numbers = TRUE) +
#   ggtitle("SHAP importance")

# Absolute average SHAP interactions (off-diagonals already multiplied by 2)
shapdfTRB <- shap_valuesTRB |> 
  sv_interaction(kind = "no")



epTRB <- data.frame(shapdfTRB) %>% 
 filter(if_any(starts_with('NC_'), ~ . > 0.001)) %>% 
  rownames_to_column() %>% 
  pivot_longer(cols = starts_with('NC_')) %>% 
  filter(value > 0.001) %>% 
  rename("SNP1" = "rowname",
         "SNP2" = "name") %>% 
  filter(SNP1 != SNP2) %>% 
  separate_wider_delim(SNP1, delim = ".1_", names = c("NC_SNP1", "POS_SNP1")) %>% 
  separate_wider_delim(POS_SNP1, delim = "_", names = c("POS_SNP1", "N1")) %>%
  mutate(NC_SNP1 = paste0(NC_SNP1, ".1")) %>% 
  separate_wider_delim(SNP2, delim = ".1_", names = c("NC_SNP2", "POS_SNP2")) %>% 
  separate_wider_delim(POS_SNP2, delim = "_", names = c("POS_SNP2", "N2")) %>%
  mutate(NC_SNP2 = paste0(NC_SNP2, ".1")) %>% 
  merge(chr, by.x='NC_SNP1', by.y='CHROM') %>% 
  rename("NumberSNP1" = "Number") %>% 
  merge(chr, by.x='NC_SNP2', by.y='CHROM') %>% 
  rename("NumberSNP2" = "Number") %>% 
  mutate(ID1 = paste0(NC_SNP1, "_", POS_SNP1)) %>% 
  mutate(ID2 = paste0(NC_SNP2, "_", POS_SNP2)) 

mTRB <- epTRB %>% 
  merge(trbXgImp, by.x = "ID1", by.y = "ID", all.x = TRUE, all.y = TRUE) #%>% 
  #select(c(ID1, ID2, NumberSNP1, NumberSNP2, value, Gain, rank))
# 
# mTRB %>% 
#   ggplot(aes(x = value, y = Gain)) + 
#   geom_point()+
#   theme_bw()

# SHAP analysis
df_explainF <- FForRF[1:53, ]

X_predF <- bake(  # Goes to xgboost:::predict.xgb.Booster()
  prep(fRecipe), 
  has_role("predictor"),
  new_data = df_explainF,
  composition = "matrix"
)

stopifnot(colnames(X_predF) %in% colnames(df_explainF))

shap_values <- extract_fit_engine(fXgFit) |> 
  shapviz(X_pred = X_predF, X = df_explainF, interactions = TRUE)

# SHAP importance
# shap_values |> 
#   sv_importance(show_numbers = TRUE) +
#   ggtitle("SHAP importance")

# Absolute average SHAP interactions (off-diagonals already multiplied by 2)
shapdf <- shap_values |> 
  sv_interaction(kind = "no")

ep <- data.frame(shapdf) %>% 
 filter(if_any(starts_with('NC_'), ~ . > 0.001)) %>% 
  rownames_to_column() %>% 
  pivot_longer(cols = starts_with('NC_')) %>% 
  filter(value > 0.001) %>% 
  rename("SNP1" = "rowname",
         "SNP2" = "name") %>% 
  filter(SNP1 != SNP2) %>% 
  separate_wider_delim(SNP1, delim = ".1_", names = c("NC_SNP1", "POS_SNP1")) %>% 
  separate_wider_delim(POS_SNP1, delim = "_", names = c("POS_SNP1", "N1")) %>%
  mutate(NC_SNP1 = paste0(NC_SNP1, ".1")) %>% 
  separate_wider_delim(SNP2, delim = ".1_", names = c("NC_SNP2", "POS_SNP2")) %>% 
  separate_wider_delim(POS_SNP2, delim = "_", names = c("POS_SNP2", "N2")) %>%
  mutate(NC_SNP2 = paste0(NC_SNP2, ".1")) %>% 
  merge(chr, by.x='NC_SNP1', by.y='CHROM') %>% 
  rename("NumberSNP1" = "Number") %>% 
  merge(chr, by.x='NC_SNP2', by.y='CHROM') %>% 
  rename("NumberSNP2" = "Number") %>% 
  mutate(ID1 = paste0(NC_SNP1, "_", POS_SNP1)) %>% 
  mutate(ID2 = paste0(NC_SNP2, "_", POS_SNP2)) 

m <- ep %>% 
  merge(fXgImp, by.x = "ID1", by.y = "ID", all.x = TRUE, all.y = TRUE)# %>% 
 # select(c(ID1, ID2, NumberSNP1, NumberSNP2, value, Gain, rank))

m %>% 
  ggplot(aes(x = NumberSNP1, y = NumberSNP2, colour = value, alpha = 0.5, size = Gain)) + 
  geom_point()+
  theme_bw()

```



# To-do and Next Steps

To do: \
- Tune RF and xgboost models\
- Finish genotyping aquaculture fish and run admixture, then model CTMax by Aquaculture Ancestry\
- Polygenic Risk Scores\
- Methylation data\

Related Projects: \
- Make code for Manhattan Plots shareable on GitHub\
- Compare GONE results for aquaculture introgressed popultaions vs. less introgressed populations

